QTL information page and tutorial

### 2023-06-01 DRAFT_QTL_tutorial.Rmd
added some code and some text to the QTL tutorial

## Tutorial
I used whole genome skim sequencing due to the ease of a multiplex library prep, making costs and wet lab time comparable to reduced representation methods

### general info on calling variants
If you have a reference genome available, you can of course just call variants against it. If not, you’ll do a de novo assembly from your sequence data (link tutorials, packages for this within and outside MarineOmics). You may also use another program for calling SNPs without a reference or enough sequence for de novo assembly.

@@

This whole idea of steps 1-3 below is to determine the genotypes of the F2 population at enough useful sites to properly represent their genotypes across recombination bins. By useful, I mean only sites that were polymorphic between the parents, and within these, only sites where the F1 inherited different alleles from each parent (heterozygous).

### Let's start by determining the useful sites for genotyping the F2s
####  Generate a genotype depth sequence file for the F1
  * Of course, this can be done by sequencing the F1 to enough coverage to confidently genotype it at many SNPs, however there is a much easier and less costly way:
    * In this case, the F2 population only inherited alleles from the F1, so if we merge all of the F2 sequence data, we're left with a genotype level sequence file of the F1!

@@

>>>>

## Tutorial
This tutorial will walk you through the essentials of the particular analysis that I did, but my hope is that it will be useful even for those who take different approaches. It is not written with reproducible data, but I have included all of the necessary code and outputs.

I used whole genome skim sequencing due to the ease of a multiplex library prep, making costs and wet lab time comparable to reduced representation methods.

### general info on calling variants
If you have a reference genome available, you can of course just call variants against it. If not, you’ll do a de novo assembly from your sequence data (link tutorials, packages for this within and outside MarineOmics). You may also use another program for calling SNPs without a reference or enough sequence for de novo assembly. I have included all script files for this analysis here. For the sake of the tutorial, code is split into chunks with its output. For a brush up on command line code, Nicolas Lu has a great tutorial in his WGS page (if this is true, link it).

@@

The whole idea of steps 1-3 below is to determine the genotypes of the F2 population at enough useful sites to properly represent their genotypes across recombination bins. By useful, I mean only sites that were polymorphic between the parents, and within these, only sites where the F1 inherited different alleles from each parent (heterozygous).

### We're going to start by defining all of our filepaths, variable names, and tools
```{bash, eval = F}
# make variables for commonly used directories
workingdir="/Users/jason/qtl.tutorial"
mkdir $workingdir

bamdir="$workingdir/f2.alignments"
mkdir $bamdir

infodir="$workingdir/info"
mkdir $infodir

genodir="$workingdir/genotype"
idallelesdir="$genodir/id.alleles"
mkdir $genodir
mkdir $idallelesdir

# file with the Aquilegia reference genome
reference="/Volumes/HD/genomes.reference/Aquilegia/V3/sequences/Aquilegia_coerulea.main_genome.scaffolds.fasta"
f1bam="$infodir/f2.merged.bam"

# on our lab computer we call samtools from it's location
samtools="/Volumes/HD/tools/samtools-1.3/samtools"

```

### To begin our genotyping, let's determine the useful sites for genotyping the F2s
####  Generate a genotype depth sequence file for the F1
  * Of course, this can be done by sequencing the F1 to enough coverage to confidently genotype it at many SNPs, however there is a much easier and less costly way:
    * In this case, the F2 population only inherited alleles from the F1, so if we merge all of the F2 sequence data, we're left with a genotype level sequence file of the F1!

```{bash, eval = F}
# we'll start in our working directory
cd $workingdir

# make a list file with all of the file paths of our .bams for samtools to merge
ls -d -1 $PWD/*.* > $infodir/f2.paths.list

# merge bam files
samtools merge $infodir/f2.merged.bam -b $infodir/f2.paths.list

# index bam files
samtools index $infodir/f2.merged.bam
```

We can view the merged, indexed .bam file of the F2s in IGV (link). Because I got ~1x coverage on ~350 individuals, this merged .bam file is essentially a 350x coverage sequence file of the F1...plenty to call genotypes :).

@@

```{bash, eval = F}
# move into the id.alleles directory
cd $idallelesdir

# run samtool mpileup to identify variant sites and create a .vcf file with those sites in our 'F1'
samtools mpileup -q 30 -B -u -v -f $reference $f1bam | bcftools call -vm -f GQ > var.f1.vcf

```
